### 2020-11-11 paper.tex
writed abstract alpha version

\documentclass{article}

\usepackage{arxiv}

\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage{hyperref}       % hyperlinks
\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography
\usepackage{lipsum}		    % Can be removed after putting your text content
\usepackage{graphicx}
\usepackage{natbib}
\usepackage{doi}

\title{A hybrid approach for news recommender system using optimization methods}

%\date{September 9, 1985}	% Here you can change the date presented in the paper title
%\date{} 					% Or removing it

\author{
    {\includegraphics[scale=0.07]{./images/spbu.png}\hspace{1mm}Alexander Smirnov}\\
	Department of Information and Analytical Systems\\
	Saint Petersburg State University\\
	Russia, Saint Petersburg\\
	\texttt{ru.alexander.smirnov@gmail.com} \\
	\And
	{\includegraphics[scale=0.07]{images/spbu.png}\hspace{1mm}Elena Mikhailova} \\
	Department of Information and Analytical Systems\\
	Saint Petersburg State University\\
	Russia, Saint Petersburg\\
	\texttt{e.mikhaylova@spbu.ru} \\
	%% \AND
	%% Coauthor \\
	%% Affiliation \\
	%% Address \\
	%% \texttt{email} \\
	%% \And
	%% Coauthor \\
	%% Affiliation \\
	%% Address \\
	%% \texttt{email} \\
	%% \And
	%% Coauthor \\
	%% Affiliation \\
	%% Address \\
	%% \texttt{email} \\
}

% Uncomment to remove the date
%\date{}

% Uncomment to override  the `A preprint' in the header
%\renewcommand{\headeright}{Technical Report}
%\renewcommand{\undertitle}{Technical Report}
\renewcommand{\shorttitle}{Recommender System}

\hypersetup{
    pdftitle={A hybrid approach for news recommender system using optimization methods},
    pdfsubject={recommender systems},
    pdfauthor={Alexander Smirnov, Elena Mikhailova},
    pdfkeywords={Recommender Systems, Optimizations},
}

\begin{document}

    \maketitle

    \begin{abstract}

        There are many approaches \citep{approaches} to build recommender system. They have their advantages and disadvantages. In this work we address the question whether it worths to use as many information as we are able to collect to build a hybrid recommender system. This work is aimed to combine multiple recommenders via optimizing their contributions to weights vectors. Results show that combining different approaches leads to rise of users' involvement.

    \end{abstract}

    \keywords{Recommender Systems \and Optimizations}

    \section{Introduction}

    \bibliographystyle{unsrtnat}
    \bibliography{references}

\end{document}

### 2020-12-02 paper/paper.tex
moved paper to separate folder

--- 2020-11-11 paper.tex
@@
 
+        Nowadays recommender systems are crucial part of every service that operates with user activity. There is so many information that user needs help to find content he wants to consume.
+
+        In this paper we found that it would be great to combine all available recommenders with different weights to negotiate their weak sides.
+
+        We use \textbf{Collaborative Filtering} and \textbf{Content-Based filtering} as main approaches.
+
+        Previously there were works related to single recommendation method, but there are not many of them that combine all approaches together and optimize them.
+
+        Vocabulary:
+
+            \begin{itemize}
+                \item Rating: expression or preference
+                \item Explicit (direct from user, e.g. rate the film)
+                \item Implicit (inferred from user activity, e.g. stop watching movie after 5 min)
+                \item Prediction: estimate of preference
+                \item Recommendation: selected items for user
+                \item Content: attributes, text, etc; everything about item
+                \item Collaborative: using data from other users
+            \end{itemize}
+
+        This work aimed to prove that hybrid methods works better than single recommenders. It will fill a gap between content creators and users so it will be easier to deliver content.
+
+        Firstly we will implement all state-of-the-art solutions. Afterwards we will combine them and optimize.
+
+        Results show that combining different approaches leads to rise of users' involvement.
+
     \bibliographystyle{unsrtnat}

### 2020-12-21 thesis/diploma.tex
added part of thesis work

\documentclass[14pt]{matmex-diploma-custom}

\usepackage[section]{placeins}

\begin{document}

\filltitle{ru}{
	chair              = {Математическое обеспечение и администрирование информационных систем\\Кафедра информационно-аналитических систем},
	title              = {Построение гибридной рекомендательной системы новостей с применением методов оптимизации},
	type               = {diploma},
	position           = {студента},
	group              = 441,
	author             = {Смирнов Александр Львович},
	supervisorPosition = {к.ф.-м.н., доц.},
	supervisor         = {Михайлова Елена Георгиевна}
}

\maketitle

\tableofcontents

\section*{Введение}

Невозможно представить современное приложение, оперирующее большими объёмами информации, без рекомендательной системы. Информации становится настолько много, что пользователи не имеют возможности найти интересный себе контент. Для решения данной проблемы применяются рекомендательные системы.

Рекомендательная система -- совокупность алгоритмов, которые принимают на вход данные об активности пользователей и информацию о рекомендуемых предметах (дата создания, источник и т.д.) и на выходе выдают список предметов, отранжированный по мере того, насколько он релевантен конкретному пользователю.

Компании IT Service требуется рекомендательная система новостей, для того, чтобы увеличить вовлечённость пользователей путём рекомендации релевантного контента.
Ключевым фактором является то, что отстутствует информация о действиях пользователей, вследствие чего не представляется возможным оценить работу алгоритмов на реальных данных.

\section{Постановка задачи}

Целью данной работы является построение и оптимизация гибридной рекомендательной системы новостей. Для её достижения необходимо решить следующие задачи:

\begin{itemize}
	\item провести обзор существующих решений
	\item реализовать каждое уникальное решение и оценить его
	\item соединить различные решения в единую систему
	\item найти оптимальный вклад каждого отдельного решения в результирующий список рекомендаций
\end{itemize}

\section{Обзор}

\subsection{Методы}

В области рекомендательных систем различают два основных подхода:

\begin{itemize}
	\item collaborative filtering
	      \begin{itemize}
		      \item рекомендации на основе действий других пользователей
	      \end{itemize}
	\item content-based filtering
	      \begin{itemize}
		      \item рекомендации на основе содержимого
	      \end{itemize}
\end{itemize}

Оба подхода обладают своими плюсами и минусами. К плюсам коллаборативной фильтрации можно отнести отсутствие необходимости знать что-либо о рекомендуемом предмете, так как рекомендации строятся на основе действий других пользователей. К минусам относится необходимость иметь данные об активности других пользователей, что называется ''холодный старт''.

В свою очередь рекомендации на основе контента не требуют данных о действиях других пользователей, но необходимы признаки, которыми обладает рекомендуемый предмет, для построения вектора предпочтений пользователя.

\subsection{Подходы}

Для коллаборативной фильтрации используют метод kNN \cite{wiki:knn} и корреляцию Пирсона для нахождения похожих векторов.

Для рекомендации на основе содержимого используются байесовский классификатор, решающие деревья, кластерный анализ.

\section{Решение}

В качестве решения предлагается комбинация подходов:

\begin{figure}[h]
	\includegraphics[width=0.5\textwidth]{./images/main.png}
	\centering
\end{figure}

Подходы:

\begin{itemize}
	\item Content-based recommender
	\item Collaborative recommender
	\item Session recommender
	      \begin{itemize}
		      \item рекомендации на основе последних просмотренных элементов
	      \end{itemize}
	\item Boost recommender
	      \begin{itemize}
		      \item рекомендации на основе информации об элементе
	      \end{itemize}
\end{itemize}

Каждый из сервисов уже был реализован, остаётся решить задачу оптимизации, то есть найти подходящие веса для каждого из подходов.

\section{Дальнейшие планы}

В начале января планируется релиз приложение и масштабная пиар-компания, вследствие чего будет доступно очень много логов пользовательской активности.

Собирая и используя эту информацию мы сможем оценить качество построенных моделей.

Также планируется проводить A / B тестирование.

\setmonofont[Mapping=tex-text]{CMU Typewriter Text}
\bibliographystyle{ugost2008ls}
\bibliography{diploma.bib}
\end{document}

### 2021-01-09 paper/paper.tex
created paper structure

--- 2020-11-11 paper.tex
@@
 \usepackage{doi}
+\usepackage{enumitem}       % itemize settings
+\usepackage{cleveref}       % reference section with capital letter
+\usepackage{xcolor}         % additional colors for hypersetup
+
+\setlist[itemize]{noitemsep, topsep=0pt} % no space between itemize
+
+\hypersetup{
+    colorlinks=true,
+    citecolor=violet,
+    linkcolor=violet,
+    filecolor=violet,      
+    urlcolor=black,
+}
 
@@
 
-        There are many approaches \citep{approaches} to build recommender system.
-They have their advantages and disadvantages.
-In this work we address the question whether it worths to use as many information as we are able to collect to build a hybrid recommender system.
-This work is aimed to combine multiple recommenders via optimizing their contributions to weights vectors.
-Results show that combining different approaches leads to rise of users' involvement.
+        Recommender system is an essential part of any social media application.
+Most recommender systems now use a hybrid approach, combining collaborative filtering, content-based filtering, and other approaches.
+Most common problems in the field of hybrid recommenders are cold start and data sparsity \citep{overview}.
+In this paper we address the abovementioned problems by proposing a hybrid weighted news recommender system which combines different approaches.
 
@@
 
-    \keywords{Recommender Systems \and Optimizations}
+    \keywords{recommender systems \and content-based recommender \and collaborative recommender \and optimizations}
 
     \section{Introduction}
+
+        Recommender system is a crucial part of every application that operates with content and user activity.
+Enormous amount of information leads to the problem that user is not able to find relevant content.
+
+        Common approaches, such as collaborative filtering, has its own problems: cold start, scalability and data sparsity.
+Content-based approaches suffer from the fact that we have to somehow represent recommended item in feature space.
+
+        To be consistent during the paper we list some domain specific vocabulary with their meanings:
+
+            \begin{itemize}
+                \item Rating: expression or preference
+
+                    \begin{itemize}
+                        \item explicit (direct from user, e.g. user rated film)
+                        \item implicit (inferred from user activity, e.g. user stopped watching movie after 5 minutes)
+                    \end{itemize}
+                \item Prediction: estimate of preference
+                \item Recommendation: selected items for user
+                \item Content: attributes, text, etc; everything about item
+            \end{itemize}
+
+        The remainder of this paper is organized as follows:
+
+            \begin{itemize}
+                \item \Cref{sec:related} describes the relevant related work
+                \item \Cref{sec:input} describes input data
+                \item \Cref{sec:overview} explains our modular design and architecture
+                \item \Cref{sec:implementation} describes the implementation of the algorithms in a real system
+                \item \Cref{sec:evaluation} provides tests and experiments validating
+our system’s results
+                \item \Cref{sec:further} explains future work
+                \item \Cref{sec:summary} presents conclusions
+            \end{itemize}
+
+    \section{Related work}
+    \label{sec:related}
+
+    \section{Input data}
+    \label{sec:input}
+
+    \section{Overview of our approach}
+    \label{sec:overview}
+
+    \section{Implementation}
+    \label{sec:implementation}
+
+    \section{Evaluation}
+    \label{sec:evaluation}
+
+    \section{Further research}
+    \label{sec:further}
+
+    \section{Summary}
+    \label{sec:summary}
+
+        Results show that combining different approaches leads to rise of users' involvement.
 

### 2021-01-09 paper/paper.tex
added input data section

@@
 \documentclass{article}
+
+\usepackage{fontspec}                % font for emoji
+\newfontfamily\DejaSans{DejaVu Sans}
 
@@
 \usepackage{natbib}
-\usepackage{doi}
+\usepackage{doi}            % hyperlink to doi.org
 \usepackage{enumitem}       % itemize settings
@@
 \usepackage{xcolor}         % additional colors for hypersetup
+\usepackage{booktabs}       % beauful tables
+\usepackage{caption}        % table caption margin
+\usepackage{placeins}       % floatbarrier
 
+\captionsetup[table]{skip=10pt}          % table caption margin
 \setlist[itemize]{noitemsep, topsep=0pt} % no space between itemize
-
 \hypersetup{
@@
 
+    As we solving domain specific task, we have domain specific data.
+
+    \subsection*{metadata}
+
+    \begin{table}[h]
+        \centering
+        % \resizebox{\textwidth}{!}{
+        \begin{tabular}{ccccc}
+            \toprule
+
+            \emph{item\_id} & \emph{date} & \emph{source\_id} \\\midrule
+
+            1 & 2021-01-08 22:08:39 & 9  \\
+            2 & 2021-01-09 10:28:58 & 5  \\
+            3 & 2021-01-09 14:20:34 & 12 \\
+            \vdots & \vdots & \vdots  \\\bottomrule
+
+         \hline
+        \end{tabular}
+        % }
+
+        \caption{news metadata}
+        \label{tab:meta}
+    \end{table}
+
+    \begin{itemize}
+        \item \emph{source\_id}: source of news item
+    \end{itemize}
+
+    \subsection*{content}
+
+    \begin{table}[h]
+        \centering
+        \resizebox{\textwidth}{!}{
+        \begin{tabular}{cll}    \toprule
+
+            \emph{item\_id} &  \emph{news title} & \emph{news content} \\\midrule
+
+        1  &  Azerbaijan denies reports on construction of Turkish air bases in the country & Information that Turkey will create air bases ...        \\
+        2  &  Durov announced the massive transition of WhatsApp users to Telegram          & Telegram developer Pavel Durov said in his channel ...   \\
+        3  &  Passenger plane that disappeared from radar crashed                           & Passenger plane taking off from Jakarta, disappeared ... \\
+        \vdots & \vdots & \vdots \\\bottomrule
+
+         \hline
+        \end{tabular}
+        }
+
+        \caption{news item}
+        \label{tab:item}
+    \end{table}
+
+    \subsection*{shows \& views}
+
+    \begin{table}[h]
+        \parbox{.45\textwidth}{
+            \centering
+            \begin{tabular}{ccc}
+                \toprule
+
+                \emph{user\_id} & \emph{item\_id} \\\midrule
+                10 & 1  \\
+                10 & 2  \\
+                23 & 1  \\
+                23 & 3  \\
+                23 & 2  \\
+                38 & 3  \\
+                38 & 1  \\
+                \vdots & \vdots  \\\bottomrule
+            \end{tabular}
+            \caption{news shows}
+            \label{tab:shows}
+        }
+        \hfill
+        \parbox{.45\textwidth}{
+            \centering
+            \begin{tabular}{ccc}
+                \toprule
+
+                \emph{user\_id} & \emph{item\_id} \\\midrule
+                10 & 1  \\
+                10 & 2  \\
+                23 & 1  \\
+                38 & 3  \\
+                \vdots & \vdots  \\\bottomrule
+            \end{tabular}
+            \caption{news views}
+            \label{tab:views}
+            }
+    \end{table}
+
+    \begin{itemize}
+        \item \emph{shows}: if \emph{item\_id} was shown to the \emph{user\_id}
+        \item \emph{views}: if \emph{item\_id} was clicked by the \emph{user\_id}
+    \end{itemize}
+
+    \newpage
+
+    \subsection*{emotions \& comments}
+
+    There is an option to react on item via leaving emotion and/or writing a comment.
+
+    \begin{table}[h]
+        \parbox{.45\textwidth}{
+            \centering
+            \begin{tabular}{ccc}
+                \toprule
+
+                \emph{user\_id} & \emph{item\_id} & \emph{emotion\_id} \\\midrule
+                10 & 1 & 1  \\
+                10 & 2 & 3        \\
+                23 & 1 & 3        \\
+                38 & 3 & 2        \\
+                \vdots & \vdots & \vdots  \\\bottomrule
+
+            \end{tabular}
+            \caption{users' emotions}
+            \label{tab:emotion}
+        }
+        \hfill
+        \parbox{.45\textwidth}{
+            \centering
+            \resizebox{0.45\textwidth}{!}{
+            \begin{tabular}{ccc}
+                \toprule
+
+                \emph{user\_id} & \emph{item\_id} & \emph{comment} \\\midrule
+                10 & 1 & that's great  \\
+                10 & 1 & wish it will continue        \\
+                23 & 2 & whatsapp is not competetive anymore        \\
+                \vdots & \vdots & \vdots  \\\bottomrule
+
+            \end{tabular}
+        }
+            \caption{users' comments}
+            \label{tab:comment}
+            }
+    \end{table}
+
+    \begin{itemize}
+        \item \emph{emotion\_id}: one of \{{\DejaSans 😄, 😲, 😞, 😡, ♥ }\}
+    \end{itemize}
+
+    % \FloatBarrier
+
+    \subsection*{users' subscriptions}
+
+    If \emph{user\_id} subscribed to the \emph{source\_id}.
+
+    \begin{table}[h]
+        \centering
+        \begin{tabular}{cc}
+            \toprule
+
+            \emph{user\_id} & \emph{source\_id} \\\midrule
+
+            10 & 9  \\
+            23 & 5  \\
+            \vdots & \vdots  \\\bottomrule
+
+        \end{tabular}
+
+        \caption{users' subscriptions}
+        \label{tab:subscriptions}
+    \end{table}
+
     \section{Overview of our approach}
     \label{sec:overview}
+
+    \subsection{Popularity-based recommendation}
+
+    We are able to aggreage number of comments and emotions to measure popularity.
+
+    For measuring news item popularity views \& show can be aggregated from tables (cite tables) into:
+
+    \begin{table}[h]
+        \centering
+        % \resizebox{\textwidth}{!}{
+        \begin{tabular}{ccccc}
+            \toprule
+
+            \emph{item\_id} & \emph{date} & \emph{source\_id} & \emph{shows}                                                               & \emph{views}                                                 \\\midrule
+
+        1 & 2021-01-08 22:08:39 & 9  & 1043 & 231        \\
+        2 & 2021-01-09 10:28:58 & 5  & 828 & 478  \\
+        3 & 2021-01-09 14:20:34 & 12 & 163 & 25 \\
+        \vdots & \vdots & \vdots & \vdots & \vdots \\\bottomrule
+
+         \hline
+        \end{tabular}
+        % }
+
+        \caption{news metadata}
+        \label{tab:meta}
+    \end{table}
 

### 2021-01-09 paper/paper.tex
added popularity based table

@@
 
-    \keywords{recommender systems \and content-based recommender \and collaborative recommender \and optimizations}
+    \keywords{hybrid recommender systems \and content-based recommender \and collaborative recommender \and optimizations}
 
@@
             \caption{news shows}
-            \label{tab:shows}
+            \label{tab:show}
         }
@@
             \caption{news views}
-            \label{tab:views}
+            \label{tab:view}
             }
@@
 
+    \subsection{Collaborative filtering recommendation}
+
+    For collaborative filtering recommendation we should have something known as user-item matrix which may be formed from user activity from tables (cite tables)
+
     \subsection{Popularity-based recommendation}
 
-    We are able to aggreage number of comments and emotions to measure popularity.
-
-    For measuring news item popularity views \& show can be aggregated from tables (cite tables) into:
+    For measuring news item popularity following data can be aggregated: \hyperref[tab:show]{shows}, \hyperref[tab:view]{views}, \hyperref[tab:emotion]{emotions}, \hyperref[tab:comment]{comments}.
 
@@
 
-            \emph{item\_id} & \emph{date} & \emph{source\_id} & \emph{shows}                                                               & \emph{views}                                                 \\\midrule
+            \emph{item\_id} & \emph{shows\_num} & \emph{views\_num} & \emph{emotions\_num} & \emph{comments\_num} \\\midrule
 
-        1 & 2021-01-08 22:08:39 & 9  & 1043 & 231        \\
-        2 & 2021-01-09 10:28:58 & 5  & 828 & 478  \\
-        3 & 2021-01-09 14:20:34 & 12 & 163 & 25 \\
+            1 &  1043 & 231 & 52 & 7  \\
+            2 &  828 & 478 & 78 & 11 \\
+            3 &  163 & 25 & 5 & 0    \\
             \vdots & \vdots & \vdots & \vdots & \vdots \\\bottomrule
@@
 
-        \caption{news metadata}
-        \label{tab:meta}
+        \caption{aggregated popularity data}
+        \label{tab:popularity}
     \end{table}

### 2021-01-12 paper/paper.tex
updated paper

@@
 \usepackage{placeins}       % floatbarrier
+% Code insertion.
+\usepackage[outputdir=build]{minted}
 
@@
 
+    According to the study \citep{progress}, deep learning techniques are not supposed to beat conceptually and computationally simpler algorithms, so we won't touch them.
+
+    Our goal is to choose optimal algorithm for each of the following tasks:
+
+    \begin{itemize}
+        \item \textbf{Collaborative filtering}: generating predictions about the interests of a user by collecting preferences or taste information from other users.
+It is based on the assumption that if a person A has the same opinion as a person B on an issue, A is more likely to have B's opinion on a different issue than that of a randomly chosen person
+        \item \textbf{Content-based filtering}:
+        \item \textbf{Session filtering}:
+        \item \textbf{Popularity filtering}:
+        \item \textbf{Demographic filtering}:
+        \item \textbf{Time-based filtering}:
+    \end{itemize}
+
+    We face the problem that millions of news are theoretically suitable for being recommended, so it is not correct to use abovementioned methods on such large corpus of data.
+Instead, as stated in \citep{youtube} paper, we want to implement \mintinline{bash}|candidate generation| $\to$ \mintinline{bash}|ranking| pipeline to reduce number of candidates.
+
+    \subsection{Collaborative filtering}
+
+    There were many studies on this topic, but paper \citep{evaluation} proves that well-tuned basic SVD++ approach beats newely presented algorithms.
+
     \section{Input data}
@@
 
-    \newpage
-
     \subsection*{emotions \& comments}
@@
 
-    \subsection{Collaborative filtering recommendation}
+    Out goal is to combine state-of-the-art approaches in recommender systems.
+
+    Our approach consists of combining several techniques:
+
+    \begin{itemize}
+        \item \textbf{Collaborative filtering}: 
+        \item \textbf{Content-based filtering}:
+        \item \textbf{Session filtering}:
+        \item \textbf{Popularity filtering}:
+        \item \textbf{Demographic filtering}:
+        \item \textbf{Time-based filtering}:
+    \end{itemize}
+
+    Solution consists of 2 parts:
+
+    \begin{itemize}
+        \item Candidate generation
+        \item Ranking
+    \end{itemize}
+
+    \subsection{Architecture overview}
+
+    \subsection{Components overview}
+
+    \subsubsection{Collaborative filtering recommendation}
+
+    We are using SVD++ algorithm.
 
@@
 
-    \subsection{Popularity-based recommendation}
+    \subsubsection{Popularity-based recommendation}
 

### 2021-01-12 paper/images/src/time_ranking.py
added images

@@ -1,0 +1,61 @@
+import numpy as np
+import matplotlib.pyplot as plt
+
+time_passed = range(72)
+
+def score(v, t, g=1.8):
+    return (v - 1) / (t + 2)**g
+
+v = 10
+g = 1.8
+scores = [score(v, ti, g) for ti in time_passed]
+plt.plot(time_passed, scores, label="$v_i$={}, $G$={},".format(v, g))
+
+v = 100
+g = 1.8
+scores = [score(v, ti, g) for ti in time_passed]
+plt.plot(time_passed, scores, label="$v_i$={}, $G$={},".format(v, g))
+
+v = 1000
+g = 1.8
+scores = [score(v, ti, g) for ti in time_passed]
+plt.plot(time_passed, scores, label="$v_i$={}, $G$={},".format(v, g))
+
+plt.ylim(0, 3)
+plt.xlim(0, 71)
+
+plt.xlabel("hours since creation")
+plt.ylabel("score")
+
+plt.legend(loc="upper right")
+plt.savefig("../time_ranking_1.png")
+# plt.show()
+
+
+plt.clf()
+
+
+v = 100
+g = 1.5
+scores = [score(v, ti, g) for ti in time_passed]
+plt.plot(time_passed, scores, label="$v_i$={}, $G$={},".format(v, g))
+
+v = 100
+g = 1.8
+scores = [score(v, ti, g) for ti in time_passed]
+plt.plot(time_passed, scores, label="$v_i$={}, $G$={},".format(v, g))
+
+v = 100
+g = 2.1
+scores = [score(v, ti, g) for ti in time_passed]
+plt.plot(time_passed, scores, label="$v_i$={}, $G$={},".format(v, g))
+
+plt.ylim(0, 3)
+plt.xlim(0, 71)
+
+plt.xlabel("hours since creation")
+plt.ylabel("score")
+
+plt.legend(loc="upper right")
+plt.savefig("../time_ranking_2.png")
+# plt.show()

### 2021-01-12 paper/paper.tex
added time-based filtering

@@
 \usepackage{caption}        % table caption margin
+\usepackage{subcaption}     % images next to each other
 \usepackage{placeins}       % floatbarrier
-% Code insertion.
-\usepackage[outputdir=build]{minted}
+\usepackage[outputdir=build]{minted} % code insertion
 
@@
     We face the problem that millions of news are theoretically suitable for being recommended, so it is not correct to use abovementioned methods on such large corpus of data.
-Instead, as stated in \citep{youtube} paper, we want to implement \mintinline{bash}|candidate generation| $\to$ \mintinline{bash}|ranking| pipeline to reduce number of candidates.
+Instead, as stated in \citep{youtube} paper, we want to implement \texttt{candidate generation} $\to$ \texttt{ranking} pipeline to reduce number of candidates.
 
@@
 
+    \newpage
     \section{Overview of our approach}
@@
 
-    Our approach consists of combining several techniques:
-
-    \begin{itemize}
-        \item \textbf{Collaborative filtering}: 
-        \item \textbf{Content-based filtering}:
-        \item \textbf{Session filtering}:
-        \item \textbf{Popularity filtering}:
-        \item \textbf{Demographic filtering}:
-        \item \textbf{Time-based filtering}:
-    \end{itemize}
-
     Solution consists of 2 parts:
@@
     \begin{itemize}
-        \item Candidate generation
-        \item Ranking
+        \item \textbf{Candidate generation}: lowering number of items to recommend.
+These candidates are intended to be generally relevant to the user with high precision.
+The candidate generation part only provides broad personalization
+        \item \textbf{Ranking}: applying state-of-the-art algorithms to rank candidates generated on previous step
     \end{itemize}
 
-    \subsection{Architecture overview}
+    Architecture provided below on \cref{fig:architecture}:
 
-    \subsection{Components overview}
+    \begin{figure}[h]
+        \centering
+        \includegraphics[width=0.8\textwidth]{./images/architechture.pdf}
+        \caption{architecture}
+        \label{fig:architecture}
+    \end{figure}
 
-    \subsubsection{Collaborative filtering recommendation}
+    Candidate generation stage not only filters items, but also attaches weight to items.
+
+    \subsection{Candidate generation components overview}
+
+    Goal of candidate generation step is to remove generally irrelevant items so furher algorithms won't suffer from the amount of data.
+
+    \subsubsection{Time-based filtering}
+
+    As we operating with news data, first filter is the time filter. This part consists of 2 steps:
+
+    \begin{itemize}
+        \item \textbf{Filtering}: remove all items which are older than 3 days
+        \item \textbf{Ranking}: attach weights to all items left from filtering
+    \end{itemize}
+
+    For ranking we will use following formula:
+
+    $$r_i = \frac{(v_i - 1)}{(t - t_i + 2)^G}$$
+
+    \begin{itemize}
+        \item $r_i$ -- score for $item_i$
+        \item $v_i$ -- number of views of $item_i$
+        \item $t$ -- time right now, $t_i$ -- time of creation of $item_i$, $t - t_i$ -- hours passed since item created
+        \item $G$ -- gravity factor
+    \end{itemize}
+
+    The score decreases as $t - t_i$ increases, meaning that older items will get lower and lower scores.
+$v_i$ gets substracted by $-1$ to negate submitter's view.
+$t - t_i$ increased by $2$ so even if $t - t_i = 0$ gravity factor $G$ will take effect.
+
+    Below are show dependencies between score and hours since creation:
+
+    \begin{figure}[h]
+      \begin{subfigure}[b]{0.5\textwidth}
+        \includegraphics[width=\textwidth]{images/time_ranking_1.png}
+        \caption{different views number}
+      \end{subfigure}
+      \hfill
+      \begin{subfigure}[b]{0.5\textwidth}
+        \includegraphics[width=\textwidth]{images/time_ranking_2.png}
+        \caption{different gravity factor}
+      \end{subfigure}
+    \caption{time-based ranking scoring}
+        \label{fig:samples}
+    \end{figure}
+
+    At the end scores are normalized by \texttt{min-max} normalization.
+
+    \subsubsection{Demographic filtering}
+
+    \subsection{Ranking components overview}
+
+    \subsubsection{Collaborative filtering}
 
@@
 
-    \subsubsection{Popularity-based recommendation}
+    \subsubsection{Popularity filtering}
 
@@
 
+    \subsubsection{Session filtering}
+
+    \subsubsection{Content-based filtering}
+
     \section{Implementation}
@@
     \label{sec:evaluation}
+
+    \subsection{Online}
+
+    \subsection{Offline}
 

### 2021-02-25 paper/paper.tex
stated evaluation part

@@
 
+    For evaluation we have information about what recommendation list was given to each user, and what is the source of the recommendation:
+
+    \begin{table}[h]
+        \centering
+        \begin{tabular}{cccc}
+            \toprule
+            \textit{user\_id} & \textit{recommendation\_list}       & \textit{content\_based\_filtering}  & \textit{collaborative\_filtering} \\
+            \midrule
+            2 & \{(2, 0.91), (1, 0.74), (3, 0.23)\} & \{(2, 0.45), (1, 0.54), (3, 0.08)\} & \{(2, 0.92), (1, 0.4), (3, 0.3)\}\\
+
+            1 & \{(3, 0.73), (1, 0.69), (2, 0.15)\} & \{(2, 0.6), (1, 0.44), (3, 0.04)\} & \{(2, 0.58), (1, 0.58), (3, 0.14)\}\\
+            \vdots & \vdots & \vdots & \vdots \\
+            \bottomrule
+        \end{tabular}%
+
+        \caption{recommendations' logs}
+        \label{tab:recommendation_logs}
+    \end{table}
+
+    \begin{itemize}
+        \item \textit{recommendation\_list}: list of recommendations that consists of pairs (\textit{item\_id}, \textit{score})
+        \item \textit{content\_based\_filtering} \& \textit{collaborative\_filtering}: sources of recommendation
+    \end{itemize}
+
+    As we use weighted sum of our recommmenders, we have unique boost values for every user:
+
+    \begin{table}[h]
+        \centering
+        \begin{tabular}{ccc}
+            \toprule
+            \textit{user\_id} & \textit{content\_based\_filtering} & \textit{collaborative\_filtering} \\
+            \midrule
+            1                 & 0.25                                  & 1                               \\
+            2                 & 1                               & 0.5                                \\
+            \vdots & \vdots & \vdots \\
+            \bottomrule
+            \end{tabular}%
+        \caption{boost values}
+        \label{tab:boost_values}
+    \end{table}
+
+    To illustrate score calculation of recommendations, take a look at 1st row of table \hyperref[tab:recommendation_logs]{recommendations' logs}.
+We see recommendation $r = \{(2, 0.91), (1, 0.74), (3, 0.23)\}$ for user $u = 2$, so we should find boost values for this particular user in table \hyperref[tab:boost_values]{boost values}: content-based filtering boost $b_{cbf}$ for $u$ is $1$, collaborative filtering boost $b_{cf}$ is $0.5$.
+So final score is calculated in the following way: $0.91 = 0.45 * b_{cbf} + 0.92 * b_{cf} = 0.45 * 1 + 0.92 * 0.5 = 0.45 + 0.46$.
+
+    Also we have information about \hyperref[tab:show]{shows} and \hyperref[tab:view]{views}, so we are able to track what item in recommendation list was clicked and what item was skipped, so depending on this information we can track if user liked item or not.
+
 \subsection{Online}

### 2021-02-25 paper/paper.tex
updated candidate generation

@@
 
-        \emph{item\_id} & \emph{date} & \emph{source\_id} \\\midrule
+        \emph{item\_id} & \emph{date} & \emph{source\_id} & \emph{category}\\\midrule
 
-        1 & 2021-01-08 22:08:39 & 9  \\
-        2 & 2021-01-09 10:28:58 & 5  \\
-        3 & 2021-01-09 14:20:34 & 12 \\
-        \vdots & \vdots & \vdots  \\\bottomrule
+        1 & 2021-01-08 22:08:39 & 9  & politics, conflicts\\
+        2 & 2021-01-09 10:28:58 & 5  & IT, social media\\
+        3 & 2021-01-09 14:20:34 & 12 & accident\\
+        \vdots & \vdots & \vdots & \vdots \\\bottomrule
 
@@
 Goal of candidate generation step is to remove generally irrelevant items so furher algorithms won't suffer from the amount of data.
+Both time-based filtering and content-based filtering have their own weight at start.
+As user more interacting with the system, content-based filtering increasing its weight.
+
+Using recommenders described below, we attach score to each item and pick top $n$ (e.g. $50 000$) items.
 
@@
 
-\subsubsection{Demographic filtering}
+\subsubsection{Content-based filtering (categories)}
+
+As was told before, content-based filtering have its own impact weight, which is small at start (we don't want to restrict user from content just because he made some random clicks), but it inscreases as user interacts with the system and we may make predictions about his categorial preferences.
 
@@
         \item \textit{recommendation\_list}: list of recommendations that consists of pairs (\textit{item\_id}, \textit{score})
-        \item \textit{content\_based\_filtering} \& \textit{collaborative\_filtering}: sources of recommendation
+        \item \textit{content\_based\_filtering \& collaborative\_filtering}: sources of recommendation
     \end{itemize}

### 2021-03-01 paper/paper.tex
added content based filtering

@@
 
-    Common approaches, such as collaborative filtering, has its own problems: cold start, scalability and data sparsity.
+    Common approaches, such as collaborative filtering, have their own problems: cold start, scalability and data sparsity.
 Content-based approaches suffer from the fact that we have to somehow represent recommended item in feature space.
@@
 It is based on the assumption that if a person A has the same opinion as a person B on an issue, A is more likely to have B's opinion on a different issue than that of a randomly chosen person
-    \item \textbf{Content-based filtering}:
-    \item \textbf{Session filtering}:
-    \item \textbf{Popularity filtering}:
-    \item \textbf{Demographic filtering}:
-    \item \textbf{Time-based filtering}:
+    \item \textbf{Content-based filtering}: is based on the assumption that people who liked items with certain attributes in the past, will like the same kind of items in the future as well.
+It makes use of item features to compare the item with user profiles and provide recommendations
+    \item \textbf{Session filtering}: is selecting cadidates based on user's activity on current session
+    \item \textbf{Popularity filtering}: uses information such as number of views, shows, comments, etc.
+    \item \textbf{Demographic filtering}: uses demographic data such as age, gender, education, etc. for identifying categories of users
+    \item \textbf{Time-based filtering}: ranking news is a way that more recent items have higher scores
 \end{itemize}
@@
 
-Candidate generation stage not only filters items, but also attaches weight to items.
+Candidate generation stage not only filters items, but also attaches weights to them.
 
@@
 \subsubsection{Content-based filtering (categories)}
+
+We are representing each news item as a vector of confidence values (how strong each item belongs to category).
+
+\begin{table}[h]
+    \centering
+    \begin{tabular}{cccccc}
+        \toprule
+        item\_id & \textit{politics} & \textit{IT} & \textit{social\_media} & $\cdots$ & \textit{confilcts} \\
+        \midrule
+        1 & 0.55 & 0 & 0 & $\cdots$ & 0.3 \\
+        2 & 0    & 0.81  & 0.62 & $\cdots$ & 0 \\
+        \vdots & \vdots & \vdots & \vdots & \vdots & \vdots\\
+
+        \bottomrule
+    \end{tabular}%
+    \caption{text vectors}
+    \label{tab:text_vectors}
+\end{table}
+
+Tagging is made in the following way: we have a dictionary of words that belongs to categories:
+
+\begin{table}[h]
+    \centering
+    \begin{tabular}{cc}
+        \toprule
+        \textit{word} & \textit{category} \\
+        \midrule
+        alexey\_nalalny & politics \\
+        crash & accident \\
+        telegeram & IT \\
+        \vdots & \vdots \\
+        \bottomrule
+    \end{tabular}%
+    \caption{words' categories}
+    \label{tab:words_categories}
+\end{table}
+
+Confidence is taken from word's tf-idf metric alongside all corpus.
+
+As user interacts with the system, we are forming his preferences vector in the following way:
+
+\begin{equation}
+    \texttt{user\_vector} \mathrel{+}= \texttt{action\_score} * \texttt{item\_vector}
+\end{equation}
+
+Where action score is taken from table:
+
+\begin{table}[h]
+    \centering
+    \begin{tabular}{cc}
+        \toprule
+        \textit{action} & \textit{score} \\
+        \midrule
+        shown but now viewed & -1 \\
+        viewed & 0.5 \\
+        emoji or comment & 0.5 \\
+        read till the end & 1 \\
+        \bottomrule
+    \end{tabular}%
+    \caption{actions' scores}
+    \label{tab:action_score}
+\end{table}
+
+As we now have both user's and items' vectors, we are able to find similarities between them:
+
+\begin{equation}
+    r_i = 1 - \cos{(\texttt{user\_vector}, \texttt{item\_vector})}
+\end{equation}
 
@@
 
-    \begin{table}[h]
+    \begin{table}[]
         \centering

### 2021-03-11 paper/paper.tex
updated paper

@@
 
-\title{A hybrid approach for news recommender system using optimization methods}
+\title{A Hybrid Approach for News Recommender System using Optimization Methods}
 
@@
 
+    Recommender systems have been studied to present items, such as movies, music, and books \citep{movies} \citep{films} \citep{books}.
+The term now has a broader connotation, describing any system that produces individualized recommendations as output or has the effect of guiding the user in a personalized way to interesting or useful objects in a large space of possible options.
+Such systems have an obvious appeal in an environment where the amount of on-line information vastly outstrips any individual’s capability to survey it.
+
+    Recommender systems are now an integral part of some e-commerce sites such as Amazon.com and CDNow \citep{commerce}.
+It is the criteria of ’individualized’ and ’interesting and useful’ that separate the recommender system from information retrieval systems or search engines \citep{coin}.
+The semantics of a search engine are ’matching’: the system is supposed to return all those items that match the query ranked by degree of match.
+Techniques such as relevance feedback enable a search engine to refine its representation of the user’s query, and represent a Simple form of recommendation.
+
+     There are three main types of recommendation methods: memorybased, model-based, and hybrid \citep{survey}.
+Memory-based methods \citep{memory} usually use similarity metrics to obtain the distance between two users or two items.
+Model-based methods use demographic, content, or aggregated information to create a model that generates recommendations.
+Hybrid \citep{combining} methods combine different types of recommenders to gain better performance.
+
     Common approaches, such as collaborative filtering, have their own problems: cold start, scalability and data sparsity.
 Content-based approaches suffer from the fact that we have to somehow represent recommended item in feature space.
+
+    So we are going to present a hybrid recommender system.
 
@@
             \item \Cref{sec:overview} explains our modular design and architecture
-            \item \Cref{sec:implementation} describes the implementation of the algorithms in a real system
+            % \item \Cref{sec:implementation} describes the implementation of the algorithms in a real system
             \item \Cref{sec:evaluation} provides tests and experiments validating
@@
     \item \textbf{Collaborative filtering}: generating predictions about the interests of a user by collecting preferences or taste information from other users.
-It is based on the assumption that if a person A has the same opinion as a person B on an issue, A is more likely to have B's opinion on a different issue than that of a randomly chosen person
+It is based on the assumption that if a person A has the same opinion as a person B on an issue, A is more likely to have B's opinion on a different issue than that of a randomly chosen person; there were many studies on this topic, but paper \citep{evaluation} proves that well-tuned basic SVD++ approach beats newely presented algorithms
     \item \textbf{Content-based filtering}: is based on the assumption that people who liked items with certain attributes in the past, will like the same kind of items in the future as well.
@@
 
+Also there is several ways \citep{survey1} to combine recommenders between each other:
+
+\begin{itemize}
+    \item \textbf{Weighted}: The scores (or votes) of several recommendation techniques are.combined together to produce a single recommendation
+    \item \textbf{Switching}: The system switches between recommendation techniques depending on the current situation
+    \item \textbf{Mixed}: Recommendations from several different recommenders are presented at the same time
+    \item \textbf{Feature combination}: Features from different recommendation data sources are thrown together into a single recommendation algorithm
+    \item \textbf{Cascade}: Features from different recommendation data sources are thrown together into a single recommendation algorithm 
+    \item \textbf{Feature augmentation}: Output from one technique is used as an input feature to another
+    \item \textbf{Meta-level}: The model learned by one recommender is used as input to another
+\end{itemize}
+
 We face the problem that millions of news are theoretically suitable for being recommended, so it is not correct to use abovementioned methods on such large corpus of data.
 Instead, as stated in \citep{youtube} paper, we want to implement \texttt{candidate generation} $\to$ \texttt{ranking} pipeline to reduce number of candidates.
-
-\subsection{Collaborative filtering}
-
-There were many studies on this topic, but paper \citep{evaluation} proves that well-tuned basic SVD++ approach beats newely presented algorithms.
 
@@
 As we solving domain specific task, we have domain specific data.
+Out data consist of information about news, so it includes following tables:
 
 \subsection*{metadata}
+
+The metadata about news.
+Here we have an \textit{item\_id} column, which stands for unique item indentifier, afterwards \textit{date}, that shows when this item was released.
+The \textit{source\_id} column stands for id of a publisher of this news item.
+\textit{category} column means category of current news item.
+This category is taken from news text by keywords.
 
@@
         \vdots & \vdots & \vdots & \vdots \\\bottomrule
-
-     \hline
     \end{tabular}
@@
 
-\begin{itemize}
-    \item \emph{source\_id}: source of news item
-\end{itemize}
+% \begin{itemize}
+%     \item \emph{source\_id}: source of news item
+% \end{itemize}
 
 \subsection*{content}
+
+Content table hold news texts in a \textit{news\_content} column.
+This is the primary information that is being used by our recommender system, because we are able to extract a lot of valuable data from text, such as topics.
 
@@
     \vdots & \vdots & \vdots \\\bottomrule
-
-     \hline
     \end{tabular}
     }
-
     \caption{news item}
@@
 \subsection*{shows \& views}
+
+As we operate with user activity we should have user logs, such as what items were clilcked at what time, so we have separate table that contains this information.
+\emph{shows} stands for \emph{item\_id} was shown to the \emph{user\_id}, \emph{views} is if \emph{item\_id} was clicked by the \emph{user\_id}.
 
@@
 
-\begin{itemize}
-    \item \emph{shows}: if \emph{item\_id} was shown to the \emph{user\_id}
-    \item \emph{views}: if \emph{item\_id} was clicked by the \emph{user\_id}
-\end{itemize}
+% \begin{itemize}
+%     \item \emph{shows}: if \emph{item\_id} was shown to the \emph{user\_id}
+%     \item \emph{views}: if \emph{item\_id} was clicked by the \emph{user\_id}
+% \end{itemize}
 
@@
 
-There is an option to react on item via leaving emotion and/or writing a comment.
+We have an explicit feedback that user may leave on a news if he wants to.
+These options includes leaving emoji, which is one of \{{\DejaSans 😄, 😲, 😞, 😡, ♥ }\} or comment, which we can analyze afterwards.
 
@@
 
-\begin{itemize}
-    \item \emph{emotion\_id}: one of \{{\DejaSans 😄, 😲, 😞, 😡, ♥ }\}
-\end{itemize}
+% \begin{itemize}
+%     \item \emph{emotion\_id}: one of \{{\DejaSans 😄, 😲, 😞, 😡, ♥ }\}
+% \end{itemize}
 
@@
 
-If \emph{user\_id} subscribed to the \emph{source\_id}.
+Each piece of news is being posted by some feed and user have an ability to subscribe to the feed.
+So it may give us useful information if user prefers content from one feed to content to another feed.
+So following table illustrates if \emph{user\_id} subscribed to the \emph{source\_id}.
 
@@
         \midrule
-        alexey\_nalalny & politics \\
+        trump & politics \\
         crash & accident \\
@@
 
-For collaborative filtering recommendation we should have something known as user-item matrix which may be formed from user activity from tables (cite tables)
+For collaborative filtering recommendation we should have something known as user-item matrix which may be formed from user activity from tables (cite tables).
+
+We use the modification of Funk MF, which factorized the user-item rating matrix as the product of two lower dimensional matrices, the first one has a row for each user, while the second has a column for each item.
+The row or column associated to a specific user or item is referred to as latent factors.
+
+$r_{u i}=\sum_{f=0}^{n } H_{u, f} W_{f, i}$
+
+While Funk MF is able to provide very good recommendation quality, its ability to use only explicit numerical ratings as user-items interactions constitutes a limitation.
+Modern day recommender systems should exploit all available interactions both explicit (e.g.
+numerical ratings) and implicit (e.g.
+likes, purchases, skipped, bookmarked).
+To this end SVD++ was designed to take into account implicit interactions as well.
+Compared to Funk MF, SVD++ takes also into account user and item bias.
+The predicted rating user $u$ will give to item $i$ is computed as:
+
+$r_{u i}=\mu+b_{i}+b_{u}+\sum_{f=0}^{n} H_{u, f} W_{f, i}$
 
@@
 
-\subsubsection{Session filtering}
+So when we apply information about news item popularity, we are able to give them scores via following algorithm:
+
+\texttt{min-max} normalizing each of \emph{shows\_num}, \emph{views\_num}, \emph{emotions\_num}, \emph{comments\_num}, then dividing by 4 (to have 1 as max after sum), and sum all of these values.
 
@@
 
-\section{Implementation}
-\label{sec:implementation}
+latent Dirichlet allocation (LDA)
+
+\subsubsection{Session filtering}
+
+same as previous
+
+% \section{Implementation}
+% \label{sec:implementation}
 
@@
 
-    \begin{table}[]
+    \begin{table}[h]
         \centering
@@
 
-\subsection{Online}
+% \subsection{Online}
 
-\subsection{Offline}
+% \subsection{Offline}
 

### 2021-03-12 paper/paper.tex
updated recommendation approaches

@@
 \subsubsection{Content-based filtering (categories)}
+\label{content}
 
@@
 
-\subsubsection{Content-based filtering}
+\subsubsection{Content-based filtering (LDA)}
 
-latent Dirichlet allocation (LDA)
+For Content-based filtering we should somehow vectorize news items and represent user preferences via these vectorized news.
+We will use Latent Dirichlet Allocation (LDA), which is a generative statistical model that allows sets of observations to be explained by unobserved groups that explain why some parts of the data are similar.
+For example, if observations are words collected into documents, it posits that each document is a mixture of a small number of topics and that each word's presence is attributable to one of the document's topics.
+
+So we are able vectorize text by the measure of how each text belongs to each category, from 0 to 1.
+For example, suppose we have 3 topic, that were extracted by LDA model:
+
+\begin{itemize}
+    \item \textbf{Topic 1:} ruble, bank, money, dollar
+    \item \textbf{Topic 2:} cooking, sugar, gordon\_ramsay
+    \item \textbf{Topic 3:} IT, smartphone, technology, hacker
+\end{itemize}
+
+And we have following text: ``Hackers use mobile emulators to steal millions of dollars''. Its vectorized form is going to be $[0.7, 0, 0.8]$.
+
+All calculations are done as was told in \ref{content}.
 
@@
 
-same as previous
+We want to instantly react on user's actions, so we applying session filtering in the following way:
+trying to find similar item to those, user have just watched. So
+
+$$r_{i}=\sum_{k=0}^{n} \textit{similarity}\{\textit{current\_item\_vector}, \textit{last\_viewed\_vector}_{k}\} \times \textit{weight}_k$$
+
+where $\textit{weight}_k$ is the weight of last viewed vector. Weight is bigger if item was seen more recently.
 

### 2021-03-12 paper/paper.tex
added summary

@@
 Techniques such as relevance feedback enable a search engine to refine its representation of the user’s query, and represent a Simple form of recommendation.
+
+    Field of news recommendation has its own specific: news are getting old really fast and they don't need to be recommended.
 
@@
 
+There is limited number of words, but during the day new words are being added and every night item vectors are being recalculated.
+
 As user interacts with the system, we are forming his preferences vector in the following way:
@@
 \end{equation}
+
+User vector has same dimension as item vector.
 
@@
 
+    So having all of this information allows us to tune the impact weight of each recommender using grid search.
+
 % \subsection{Online}
@@
 
-\section{Further research}
-\label{sec:further}
+% \section{Further research}
+% \label{sec:further}
 
@@
 
+Recommendation system has been widely used in different areas.
+Collaborative filtering focuses on rating, ignoring the features of items itself.
+In order to better evaluate customer preference on news, we use LDA model to
+calculate customer preference on news topics.
+
+In order to forecast rating on news, we take similarity of customers and correlation between customers and news into consideration.
+Experiment shows that our hybrid recommendation method based on features performances better in our social media app.
+
+What we contribute is we proposed a new hybrid recommendation method based on features to improve the performance.
+
 Results show that combining different approaches leads to rise of users' involvement.
+
+We use the average method to set the weight to adjust the predicted rating in this article, whose rationality needs to be further improved.
+Moreover, our method has yet to be tested on other data sets for its performance.
 

### 2021-03-12 paper/paper.tex
finished paper

@@
 	Saint Petersburg State University\\
+    ITMO university\\
 	Russia, Saint Petersburg\\
@@
             \item \Cref{sec:related} describes the relevant related work
+            \item \Cref{sec:overview} contains overview of our approch
             \item \Cref{sec:input} describes input data
-            \item \Cref{sec:overview} explains our modular design and architecture
+            \item \Cref{sec:approach} explains our modular design and architecture
             % \item \Cref{sec:implementation} describes the implementation of the algorithms in a real system
@@
 our system’s results
-            \item \Cref{sec:further} explains future work
             \item \Cref{sec:summary} presents conclusions
@@
 Instead, as stated in \citep{youtube} paper, we want to implement \texttt{candidate generation} $\to$ \texttt{ranking} pipeline to reduce number of candidates.
+
+\section{Overview of our approach}
+\label{sec:overview}
+
+Out goal is to combine state-of-the-art approaches in recommender systems.
+
+Solution consists of 2 parts:
+
+\begin{itemize}
+    \item \textbf{Candidate generation}: lowering number of items to recommend.
+These candidates are intended to be generally relevant to the user with high precision.
+The candidate generation part only provides broad personalization
+    \item \textbf{Ranking}: applying state-of-the-art algorithms to rank candidates generated on previous step
+\end{itemize}
+
+Architecture provided below on \cref{fig:architecture}:
+
+\begin{figure}[h]
+    \centering
+    \includegraphics[width=0.8\textwidth]{./images/architechture.pdf}
+    \caption{architecture}
+    \label{fig:architecture}
+\end{figure}
 
@@
 \newpage
-\section{Overview of our approach}
-\label{sec:overview}
+\section{Our approach}
+\label{sec:approach}
 
-Out goal is to combine state-of-the-art approaches in recommender systems.
-
-Solution consists of 2 parts:
-
-\begin{itemize}
-    \item \textbf{Candidate generation}: lowering number of items to recommend.
-These candidates are intended to be generally relevant to the user with high precision.
-The candidate generation part only provides broad personalization
-    \item \textbf{Ranking}: applying state-of-the-art algorithms to rank candidates generated on previous step
-\end{itemize}
-
-Architecture provided below on \cref{fig:architecture}:
-
-\begin{figure}[h]
-    \centering
-    \includegraphics[width=0.8\textwidth]{./images/architechture.pdf}
-    \caption{architecture}
-    \label{fig:architecture}
-\end{figure}
+Lets describe in detail candidate generation ranking pipeline.
 
@@
 
-$$r_i = \frac{(v_i - 1)}{(t - t_i + 2)^G}$$
+\begin{equation}
+    r_i = \frac{(v_i - 1)}{(t - t_i + 2)^G}
+\end{equation}
 
@@
 $t - t_i$ increased by $2$ so even if $t - t_i = 0$ gravity factor $G$ will take effect.
+By default we set $G=1.8$ and we increase it if we want to prioritize newer news over old and vice versa.
 
@@
 
-$r_{u i}=\sum_{f=0}^{n } H_{u, f} W_{f, i}$
+\begin{equation}
+    r_{u i}=\sum_{f=0}^{n } H_{u, f} W_{f, i}
+\end{equation}
 
@@
 
-$r_{u i}=\mu+b_{i}+b_{u}+\sum_{f=0}^{n} H_{u, f} W_{f, i}$
+\begin{equation}
+    r_{u i}=\mu+b_{i}+b_{u}+\sum_{f=0}^{n} H_{u, f} W_{f, i}
+\end{equation}
 
@@
 
+\begin{equation}
+    r_i = \frac{\emph{shows\_num\_norm}}{4} + \frac{\emph{views\_num\_norm}}{4} + \frac{\emph{emotions\_num\_norm}}{4} + \frac{\emph{comments\_num\_norm}}{4} 
+\end{equation}
+
 \subsubsection{Content-based filtering (LDA)}
@@
 
-\begin{itemize}
-    \item \textbf{Topic 1:} ruble, bank, money, dollar
-    \item \textbf{Topic 2:} cooking, sugar, gordon\_ramsay
-    \item \textbf{Topic 3:} IT, smartphone, technology, hacker
-\end{itemize}
+\begin{table}[h]
+\centering
+% \resizebox{\textwidth}{!}{%
+\begin{tabular}{ccc}
+    \toprule
+\textit{topic} & \textit{word} & \textit{score} \\
+    \midrule
+1 & dollar & 0.3 \\
+1 & bank & 0.2 \\
+1 & money & 0.15 \\
+2 & sugar & 0.4 \\
+2 & cooking & 0.3 \\
+3 & IT & 0.25 \\
+3 & hacker & 0.2 \\
+\bottomrule
+\end{tabular}%
+% }
+\caption{words' topics}
+\label{tab:words_topics}
+\end{table}
 
-And we have following text: ``Hackers use mobile emulators to steal millions of dollars''. Its vectorized form is going to be $[0.7, 0, 0.8]$.
+And we have following text: ``Hackers use mobile emulators to steal millions of dollars''. Its vectorized form is going to be $[0.3, 0, 0.2]$.
 
@@
 
-$$r_{i}=\sum_{k=0}^{n} \textit{similarity}\{\textit{current\_item\_vector}, \textit{last\_viewed\_vector}_{k}\} \times \textit{weight}_k$$
+\begin{equation}
+    r_{i}=\sum_{k=0}^{n} \textit{similarity}\{\textit{current\_item\_vector}, \textit{last\_viewed\_vector}_{k}\} \times \textit{weight}_k
+\end{equation}
 
@@
     Also we have information about \hyperref[tab:show]{shows} and \hyperref[tab:view]{views}, so we are able to track what item in recommendation list was clicked and what item was skipped, so depending on this information we can track if user liked item or not.
+For instance, let's consider situation that user clicked on 4th recommendation from recommendation list.
+That means that first 3 recommedations are bad, and we should penalty recommenders that prioritized these items, considering the fact that 1st item is the worst:
+
+\begin{equation}
+    \textit{penalty} = \frac{1}{\textit{num\_of\_item\_in\_rank}}
+\end{equation}
 
@@
 Results show that combining different approaches leads to rise of users' involvement.
+Firstly we had only a collaborative recommender and user averagely spend on news tab around 2 minutes.
+Now when we applied hybrid recommender users spend in average 5 minutes on news tab.
 

### 2021-05-27 paper/paper.tex
upd paper

@@
             \item \Cref{sec:summary} presents conclusions
+
         \end{itemize}
@@
 
-As user interacts with the system, we are forming his preferences vector in the following way:
-
-\begin{equation}
-    \texttt{user\_vector} \mathrel{+}= \texttt{action\_score} * \texttt{item\_vector}
-\end{equation}
-
-User vector has same dimension as item vector.
-
-Where action score is taken from table:
+We have actions while user interacts with the system and weights for items for each interaction (\cref{tab:action_score}).
 
@@
         \midrule
-        shown but now viewed & -1 \\
-        viewed & 0.5 \\
-        emoji or comment & 0.5 \\
+        shown & -1 \\
+        viewed & 2 \\
+        emoji or comment & 1 \\
         read till the end & 1 \\
@@
     \end{tabular}%
-    \caption{actions' scores}
+    \caption{actions' values}
     \label{tab:action_score}
 \end{table}
+
+All data is aggregated and each news item have an action score depending on what user had done with it.
+For instance, suppose user had viewed and item and left an emoji.
+Total score will be: $-1$ (user was shown an item) $+2$ (user viewed an item) $+1$ (user left an emoji) $=2$.
+
+As user interacts with the system, we are forming his preferences vector in the following way:
+
+\begin{equation}
+    u_i = \sum_{j=1}^{n} v_j \times s_j
+\end{equation}
+
+\begin{itemize}
+    \item $u_i$ -- vector of $user_i$
+    \item $v_j$ -- vector of $item_j$
+    \item $s_j$ -- action value of $item_j$
+\end{itemize}
+
+User vector has same dimension as item vector.
 
@@
 \begin{equation}
-    r_i = 1 - \cos{(\texttt{user\_vector}, \texttt{item\_vector})}
+    r_{ui}=\cos (\theta)=\frac{\mathbf{u} \cdot \mathbf{v}}{\|\mathbf{u}\|\|\mathbf{v}\|}=\frac{\sum_{i=1}^{n} u_{i} v_{i}}{\sqrt{\sum_{i=1}^{n} u_{i}^{2}} \sqrt{\sum_{i=1}^{n} v_{i}^{2}}}
 \end{equation}
+
+\begin{itemize}
+    \item $r_{ui}$ -- rating from $user$ to $item$
+    \item $u$ -- vector of $user$
+    \item $v$ -- vector of $item$
+\end{itemize}
 
@@
 
+\subsection{Combining}
+
+    So having all of this information allows us to tune the impact weight of each recommender using grid search. So we are optimizing weight that are making impact at each recommender.
+
+    The metric we are optimizing is MAP@20 that stands for Mean Average Precision.
+
 % \section{Implementation}
@@
 \label{sec:evaluation}
+
+\subsection{Overview}
 
@@
 
-    So having all of this information allows us to tune the impact weight of each recommender using grid search.
+\subsection{Metrics}
+
+We are using three metrics to measure the quality of our recommenders:
+
+    \begin{itemize}
+        \item RMSE
+        \item NDCG
+        \item MAP@20
+    \end{itemize}
 

### 2021-05-27 thesis/diploma.tex
increased image size
updated thesis work

@@
 
+\usepackage{fontspec}                % font for emoji
+\newfontfamily\DejaSans{DejaVu Sans}
+
 \usepackage[section]{placeins}
+\usepackage{booktabs}       % professional-quality tables
+\usepackage{cleveref}       % reference section with capital letter
+\usepackage{natbib}
+\usepackage{caption}        % table caption margin
+\usepackage{subcaption}     % images next to each other
+\usepackage{graphicx}
 
@@
 	supervisorPosition = {к.ф.-м.н., доц.},
-	supervisor         = {Михайлова Елена Георгиевна}
+	supervisor         = {Михайлова Елена Георгиевна},
+	reviewerPosition   = {руководитель отдела инженерии},
+    reviewer           = {Осипов Е.В.}
+
+}
+
+\filltitle{en}{
+	chair              = {Software and Administration of Information Systems\\Department of Information and Analytical Systems},
+	title              = {A Hybrid Approach for News Recommender System using Optimization Methods},
+	type               = {diploma},
+	position           = {student},
+	group              = 441,
+	author             = {Alexander Smirnov},
+	supervisorPosition = {professor},
+	supervisor         = {Elena Mikhaylova},
+	reviewerPosition   = {head of engineering department},
+    reviewer           = {Evgenii Osipov},
 }
@@
 
-\section*{Введение}
+\section{Введение}
 
-Невозможно представить современное приложение, оперирующее большими объёмами информации, без рекомендательной системы.
-Информации становится настолько много, что пользователи не имеют возможности найти интересный себе контент.
-Для решения данной проблемы применяются рекомендательные системы.
+Рекомендательная система является важной частью каждого приложения, в котором содержится большое количество контента и действий пользователей.
+Огромный объем информации приводит к тому, что пользователь не может найти релевантный для себя контент.
 
-Рекомендательная система -- совокупность алгоритмов, которые принимают на вход данные об активности пользователей и информацию о рекомендуемых предметах (дата создания, источник и т.д.) и на выходе выдают список предметов, отранжированный по мере того, насколько он релевантен конкретному пользователю.
+    Рекомендательные системы были испольюзуются для рекомендаций фильмов, музыки и книг \citep{movies} \citep{films} \citep{books}.
+Рекомендательной системой называется любая система, которая выдает персонализированные рекомендации или направляет пользователя к интересным или полезным объектам в большом пространстве возможных вариантов.
+Такие системы имеют очевидное преимущество в среде, где объем информации значительно превосходит возможности человека ее исследовать.
 
-Компании IT Service требуется рекомендательная система новостей, для того, чтобы увеличить вовлечённость пользователей путём рекомендации релевантного контента.
-Ключевым фактором является то, что отстутствует информация о действиях пользователей, вследствие чего не представляется возможным оценить работу алгоритмов на реальных данных.
+    Системы рекомендаций теперь являются неотъемлемой частью некоторых сайтов электронной коммерции, таких как Amazon.com и CDNow \citep{commerce}.
+Критерии «индивидуальности» и «интересности и полезности», отделяют рекомендательную систему от информационно-поисковых систем \citep{coin}.
+Семантика поисковой системы - «соответствие»: система должна возвращать все те элементы, которые соответствуют запросу, ранжированные по степени релевантности.
+Такие методы, как обратная связь по релевантности, позволяют поисковой системе уточнить представление запроса пользователя и представляют собой простую форму рекомендации.
 
-\section{Постановка задачи}
+    Сфера рекомендаций новостей имеет свою специфику: новости быстро стареют и это нужно учитывать.
 
-Целью данной работы является построение и оптимизация гибридной рекомендательной системы новостей. Для её достижения необходимо решить следующие задачи:
+     Существует три основных типа рекомендаций: на основе памяти, на основе модели и гибридные \citep{survey}.
+Методы на основе памяти \citep{memory} обычно используют метрики сходства для определения расстояния между двумя пользователями или двумя элементами.
+Методы, основанные на моделях, используют демографическую, информационную или агрегированную информацию для создания модели, которая генерирует рекомендации.
+Гибридные методы \citep{combining} комбинируют различные типы рекомендателей для повышения качества рекомендаций.
+
+    Общие подходы, такие как коллаборативная фильтрация, имеют свои проблемы: холодный старт, масштабируемость и разреженность данных.
+Контентные подходы страдают от того факта, что мы должны каким-то образом представлять рекомендуемый элемент в пространстве признаков.
+
+    В этой статье описана гибридная рекомендательная система.
+
+    Для единообразия в ходе работы мы перечисляем некоторые термины с их значениями:
 
         \begin{itemize}
-	\item провести обзор существующих решений
-	\item реализовать каждое уникальное решение и оценить его
-	\item соединить различные решения в единую систему
-	\item найти оптимальный вклад каждого отдельного решения в результирующий список рекомендаций
+            \item Рейтинг: скалярное значение, отражающее степень релевантности
+                \begin {itemize}
+                    \item явный (напрямую от пользователя, например, фильм с рейтингом пользователей)
+                    \item неявный (выводится из активности пользователя, например, пользователь перестал смотреть фильм через 5 минут)
+                \end {itemize}
+            \item Прогноз: оценка рейтинга
+            \item Рекомендация: выбранные ранжированные элементы для пользователя
+            \item Содержимое: атрибуты, текст и т. д .; все о рекомендуемом предмете
+        \end {itemize}
+
+    Остальная часть диплома организована следующим образом:
+
+        \begin {itemize}
+            \item \Cref{sec:related} описывает связанные работы
+            \item \Cref{sec:overview} содержит обзор нашего подхода
+            \item \Cref{sec:input} описывает входные данные
+            \item \Cref{sec:approach} объясняет наш подход
+            % \ item \ Cref {sec: implementation} описывает реализацию алгоритмов в реальной системе.
+            \item \Cref{sec:evaluation} предоставляет тесты и эксперименты, подтверждающие
+результаты нашей системы
+            \item \Cref{sec:summary} представляет выводы
+        \end {itemize}
+
+\section {Связанные работы}
+\label {sec:related}
+
+Согласно исследованию \citep {progress}, методы глубокого обучения не должны превосходить концептуально и вычислительно более простые алгоритмы, поэтому мы не будем их рассматривать.
+
+Наша цель - выбрать оптимальный алгоритм для каждой из следующих задач, чтобы впоследствии комбинировать их:
+
+\begin {itemize}
+    \item \textbf {Коллаборативная фильтрация}: создание прогнозов об интересах пользователя путем сбора информации о предпочтениях или вкусовых предпочтениях других пользователей.
+Он основан на предположении, что если человек A придерживается того же мнения, что и человек B, по какой-либо проблеме, A с большей вероятностью будет иметь мнение B по другому вопросу, чем мнение случайно выбранного человека; по этой теме было проведено много исследований, но статья \citep{evaluation} доказывает, что хорошо настроенный базовый подход SVD ++ превосходит недавно представленные алгоритмы глубокого обучения.
+    \item \textbf {Фильтрация на основе содержимого}: основана на предположении, что людям, которым нравились новости с определенными содержимым в прошлом, понравятся такие же элементы и в будущем.
+Из истории собирается профиль предпочтений пользователя, который впоследствии сравнивается с кандидатами на рекомендации.
+    \item \textbf {Фильтрация на основе текущей сессии}: рекомендует на основе активности пользователя в текущем сеансе.
+    \item \textbf {Фильтрация по популярности}: использует такую информацию, как количество просмотров, показов, комментариев и т. д.
+    \item \textbf {Демографическая фильтрация}: использует демографические данные, такие как возраст, пол, образование и т. д., для определения категорий пользователей.
+    \item \textbf {Фильтрация на основе времени}: это способ, при котором более свежие элементы получают более высокие оценки
+\end {itemize}
+
+Также есть несколько способов \citep {survey1} комбинировать рекомендатели между собой:
+
+\begin {itemize}
+    \item \textbf {Взвешено}: оценки (или голоса) нескольких методов рекомендаций объединяются для получения единой рекомендации.
+    \item \textbf {Переключаясь}: система переключается между методами рекомендаций в зависимости от текущей ситуации.
+    \item \textbf {Смешанный}: рекомендации от нескольких разных рекомендателей отображаются одновременно.
+    \item \textbf {Комбинированный}: функции из разных источников рекомендаций объединены в единый алгоритм рекомендаций.
+    \item \textbf {Стекинг}: выходные данные одного метода используются как входные данные для другого.
+
+\end {itemize}
+
+Мы сталкиваемся с проблемой, что миллионы новостей теоретически подходят для рекомендации, поэтому использовать вышеупомянутые методы на таком большом массиве данных некорректно.
+Вместо этого, как указано в статье \citep {youtube}, мы хотим реализовать конвейер \texttt {генерация кандидатов} $ \to $ \texttt {ранжирование}, чтобы уменьшить количество кандидатов.
+
+\section{Обзор решения}
+\label{sec:overview}
+
+Нашей задачей является скомбинировать лучшие подходы в области рекомендательных систем.
+Решение состоит из двух частей:
+
+\begin{itemize}
+    \item \textbf{Генерация кандидатов}: уменьшение количества рекомендуемых новостей. Эти кандидаты предназначены для фильтрации совсем не релевантного (напр., устаревшего) контента.
+    \item \textbf{Ранжирование}: применяем лучшие в своей области алгоритмы для ранжирования оставшихся новостей
 \end{itemize}
 
-\section{Обзор}
+Архитектура представлена на Рис \cref{fig:architecture}:
 
-\subsection{Методы}
+\begin{figure}[h]
+    \centering
+    \includegraphics[width=0.8\textwidth]{./images/architechture.pdf}
+    \caption{архитектура}
+    \label{fig:architecture}
+\end{figure}
 
-В области рекомендательных систем различают два основных подхода:
+\section{Входные данные}
+\label{sec:input}
+
+Поскольку мы решаем предметную задачу, у нас есть данные о предметной области.
+Наши данные состоят из информации о новостях и включают в себя следующие таблицы:
+
+\subsection*{метаданные}
+
+Метаданные о новостях.
+Здесь у нас есть столбец \textit {item\_id}, который обозначает уникальный идентификатор элемента, затем \textit{date}, который показывает, когда этот элемент был выпущен.
+Столбец \textit {source\_id} обозначает идентификатор издателя этой новости.
+Столбец \textit {category} означает категорию текущей новости.
+Эта категория берется из текста новости по ключевым словам.
+
+\begin{table}[h]
+    \centering
+    % \resizebox{\textwidth}{!}{
+    \begin{tabular}{ccccc}
+        \toprule
+
+        \emph{item\_id} & \emph{date} & \emph{source\_id} & \emph{category}\\\midrule
+
+        1 & 2021-01-08 22:08:39 & 9  & politics, conflicts\\
+        2 & 2021-01-09 10:28:58 & 5  & IT, social media\\
+        3 & 2021-01-09 14:20:34 & 12 & accident\\
+        \vdots & \vdots & \vdots & \vdots \\\bottomrule
+    \end{tabular}
+    % }
+
+    \caption{метаданные}
+    \label{tab:meta}
+\end{table}
+
+% \begin{itemize}
+%     \item \emph{source\_id}: source of news item
+% \end{itemize}
+
+\subsection*{содержимое}
+
+В таблице содержимого содержатся тексты новостей в столбце \\
+\textit {news\_content}. Это основная информация, которая используется нашей системой рекомендаций, потому что мы можем извлекать много ценных данных из текста, такие как темы.
+
+\begin{table}[h]
+    \centering
+    \resizebox{\textwidth}{!}{
+    \begin{tabular}{cll}    \toprule
+
+        \emph{item\_id} &  \emph{news title} & \emph{news content} \\\midrule
+
+    1  &  Azerbaijan denies reports on construction of Turkish air bases in the country & Information that Turkey will create air bases ...        \\
+    2  &  Durov announced the massive transition of WhatsApp users to Telegram          & Telegram developer Pavel Durov said in his channel ...   \\
+    3  &  Passenger plane that disappeared from radar crashed                           & Passenger plane taking off from Jakarta, disappeared ... \\
+    \vdots & \vdots & \vdots \\\bottomrule
+    \end{tabular}
+    }
+    \caption{содержимое}
+    \label{tab:item}
+\end{table}
+
+\subsection*{показы \& просмотры}
+
+Поскольку мы работаем с действиями пользователей, у нас должны быть журналы логов, например, какие элементы были нажаты и в какое время, поэтому у нас есть отдельная таблица, содержащая эту информацию.
+\emph {shows} означает, что \emph {item\_id} был показан \emph {user\_id}, \emph {views} - если \emph {item\_id} был нажат пользователем \emph {user\_id }.
+
+\begin{table}[h]
+    \parbox{.45\textwidth}{
+        \centering
+        \begin{tabular}{ccc}
+            \toprule
+
+            \emph{user\_id} & \emph{item\_id} \\\midrule
+            10 & 1  \\
+            10 & 2  \\
+            23 & 1  \\
+            23 & 3  \\
+            23 & 2  \\
+            38 & 3  \\
+            38 & 1  \\
+            \vdots & \vdots  \\\bottomrule
+        \end{tabular}
+        \caption{показы}
+        \label{tab:show}
+    }
+    \hfill
+    \parbox{.45\textwidth}{
+        \centering
+        \begin{tabular}{ccc}
+            \toprule
+
+            \emph{user\_id} & \emph{item\_id} \\\midrule
+            10 & 1  \\
+            10 & 2  \\
+            23 & 1  \\
+            38 & 3  \\
+            \vdots & \vdots  \\\bottomrule
+        \end{tabular}
+        \caption{просмотры}
+        \label{tab:view}
+        }
+\end{table}
+
+% \begin{itemize}
+%     \item \emph{shows}: if \emph{item\_id} was shown to the \emph{user\_id}
+%     \item \emph{views}: if \emph{item\_id} was clicked by the \emph{user\_id}
+% \end{itemize}
+
+\subsection*{эмоции \& комментарии}
+
+Пользователь может оставлять явную обратную связь на новости, которые он смотрит. В частности, можно оставить emoji \{{\DejaSans 😄, 😲, 😞, 😡, ♥ }\} или комментарий, которые мы анализируем впоследствие.
+
+\begin{table}[h]
+    \parbox{.45\textwidth}{
+        \centering
+        \begin{tabular}{ccc}
+            \toprule
+
+            \emph{user\_id} & \emph{item\_id} & \emph{emotion\_id} \\\midrule
+            10 & 1 & 1  \\
+            10 & 2 & 3        \\
+            23 & 1 & 3        \\
+            38 & 3 & 2        \\
+            \vdots & \vdots & \vdots  \\\bottomrule
+
+        \end{tabular}
+        \caption{эмоции}
+        \label{tab:emotion}
+    }
+    \hfill
+    \parbox{.45\textwidth}{
+        \centering
+        \resizebox{0.45\textwidth}{!}{
+        \begin{tabular}{ccc}
+            \toprule
+
+            \emph{user\_id} & \emph{item\_id} & \emph{comment} \\\midrule
+            10 & 1 & that's great  \\
+            10 & 1 & wish it will continue        \\
+            23 & 2 & whatsapp is not competetive anymore        \\
+            \vdots & \vdots & \vdots  \\\bottomrule
+
+        \end{tabular}
+    }
+        \caption{комментарии}
+        \label{tab:comment}
+        }
+\end{table}
+
+% \begin{itemize}
+%     \item \emph{emotion\_id}: one of \{{\DejaSans 😄, 😲, 😞, 😡, ♥ }\}
+% \end{itemize}
+
+% \FloatBarrier
+
+\subsection*{подписки на ленты}
+
+Каждая новость публикуется через какую-либо ленту, и у пользователя есть возможность подписаться на эту ленту.
+Таким образом, это может дать нам полезную информацию, если пользователь предпочитает контент из одной ленты контенту из другой ленты.
+Итак, следующая таблица показывает, подписался ли \emph {user\_id} на \emph{feed\_id}.
+
+\begin{table}[h]
+    \centering
+    \begin{tabular}{cc}
+        \toprule
+
+        \emph{user\_id} & \emph{source\_id} \\\midrule
+
+        10 & 9  \\
+        23 & 5  \\
+        \vdots & \vdots  \\\bottomrule
+
+    \end{tabular}
+
+    \caption{факт подписки}
+    \label{tab:subscriptions}
+\end{table}
+
+\newpage
+\section{Наш подход}
+\label{sec:approach}
+
+Давайте подробно опишем конвейер ранжирования кандидатов.
+
+Этап генерации кандидатов не только фильтрует элементы, но и присваивает им веса.
+
+\subsction {Обзор компонентов генерации кандидатов}
+
+Цель этапа генерации кандидатов - удалить, как правило, нерелевантные элементы, чтобы дополнительные алгоритмы не пострадали от объема данных.
+И фильтрация на основе времени, и фильтрация на основе содержимого имеют свой собственный вес.
+По мере того, как пользователь все больше взаимодействует с системой, фильтрация на основе содержимого увеличивает свой вес.
+
+Используя рекомендации, описанные ниже, мы присваиваем баллы каждому элементу и выбираем самые популярные элементы стоимостью $ n $ (например, $ 50 000).
+
+\subsubsection {Фильтрация по времени}
+
+Поскольку мы работаем с данными новостей, первый фильтр - это временной фильтр. Эта часть состоит из 2 шагов:
+
+\begin {itemize}
+    \item \textbf {Фильтрация}: удалить все элементы старше 3-х дней
+    \item \textbf {Рейтинг}: присвоить вес всем элементам, оставшимся после фильтрации
+\end {itemize}
+
+Для ранжирования мы будем использовать следующую формулу:
+
+\begin {equation}
+    r_i = \frac {(v_i - 1)} {(t - t_i + 2) ^ G}
+\end {equation}
+
+\begin {itemize}
+    \item $ r_i $ - счет за $ item_i $
+    \item $ v_i $ - количество просмотров $ item_i $
+    \item $ t $ - время прямо сейчас, $ t_i $ - время создания $ item_i $, $ t - t_i $ - часы, прошедшие с момента создания элемента
+    \item $ G $ - фактор гравитации
+\end {itemize}
+
+Оценка уменьшается по мере увеличения $ t - t_i $, что означает, что более старые элементы будут получать все более низкие оценки.
+$ v_i $ вычитается на $ -1 $, чтобы отрицать просмотр создателя новости.
+$ t - t_i $ увеличивается на $ 2 $, поэтому, даже если $ t - t_i = 0$, фактор силы тяжести $ G $ будет действовать.
+По умолчанию мы устанавливаем $ G = 1.8 $ и увеличиваем его, если хотим отдавать приоритет новым новостям над старыми и наоборот.
+
+Ниже показаны зависимости между счетом и часами с момента создания:
+
+\begin{figure}[h]
+  \begin{subfigure}[b]{0.5\textwidth}
+    \includegraphics[width=\textwidth]{images/time_ranking_1.png}
+    \caption{разное количество просмотров}
+  \end{subfigure}
+  \hfill
+  \begin{subfigure}[b]{0.5\textwidth}
+    \includegraphics[width=\textwidth]{images/time_ranking_2.png}
+    \caption{разный фактор гравитации}
+  \end{subfigure}
+\caption{ранжирование по времени}
+    \label{fig:samples}
+\end{figure}
+
+В конце значиния нормализуются с помощью \texttt{min-max} нормализации.
+
+\subsubsection{Фильтрация на основе категорий}
+\label{content}
+
+Мы представляем каждую новость как вектор значений достоверности (насколько сильно каждая новость относится к категории).
+
+\begin{table}[h]
+    \centering
+    \begin{tabular}{cccccc}
+        \toprule
+        item\_id & \textit{politics} & \textit{IT} & \textit{social\_media} & $\cdots$ & \textit{confilcts} \\
+        \midrule
+        1 & 0.55 & 0 & 0 & $\cdots$ & 0.3 \\
+        2 & 0    & 0.81  & 0.62 & $\cdots$ & 0 \\
+        \vdots & \vdots & \vdots & \vdots & \vdots & \vdots\\
+
+        \bottomrule
+    \end{tabular}
+    \caption{тексовые вектора}
+    \label{tab:text_vectors}
+\end{table}
+
+Тегирование происходит следующим образом: составляем словарь принадлежности слов к категориям:
+
+\begin{table}[h]
+    \centering
+    \begin{tabular}{cc}
+        \toprule
+        \textit{word} & \textit{category} \\
+        \midrule
+        trump & politics \\
+        crash & accident \\
+        telegeram & IT \\
+        \vdots & \vdots \\
+        \bottomrule
+    \end{tabular}%
+    \caption{категории слов}
+    \label{tab:words_categories}
+\end{table}
+
+Уверенность берется из метрики слова tf-idf вместе со всем набором документов.
+
+Количество слов ограничено, но в течение дня добавляются новые слова, а каждую ночь пересчитываются вектора элементов.
+
+У нас есть действия, пока пользователь взаимодействует с системой, и веса для элементов для каждого взаимодействия (\cref {tab:action_score}).
+
+\begin{table}[h]
+    \centering
+    \begin{tabular}{cc}
+        \toprule
+        \textit{action} & \textit{score} \\
+        \midrule
+        shown & -1 \\
+        viewed & 2 \\
+        emoji or comment & 1 \\
+        read till the end & 1 \\
+        \bottomrule
+    \end{tabular}%
+    \caption{веса взаимодействий}
+    \label{tab:action_score}
+\end{table}
+
+Все данные агрегированы, и каждая новость имеет оценку действия в зависимости от того, что с ней сделал пользователь.
+Предположим, что пользователь просмотрел и оставил смайлик.
+Общий балл составит: $ -1 $ (пользователю был показан элемент) $ + 2 $ (пользователь просмотрел элемент) $ + 1 $ (пользователь оставил смайлик) $ = 2 $.
+
+По мере взаимодействия пользователя с системой мы формируем его вектор предпочтений следующим образом:
+
+\begin{equation}
+    u_i = \sum_{j=1}^{n} v_j \times s_j
+\end{equation}
 
 \begin{itemize}
-	\item collaborative filtering
-	      \begin{itemize}
-		      \item рекомендации на основе действий других пользователей
-	      \end{itemize}
-	\item content-based filtering
-	      \begin{itemize}
-		      \item рекомендации на основе содержимого
-	      \end{itemize}
+    \item $u_i$ -- вектор $user_i$
+    \item $v_j$ -- вектор $item_j$
+    \item $s_j$ -- вес взаимодействия с $item_j$
 \end{itemize}
 
-Оба подхода обладают своими плюсами и минусами.
-К плюсам коллаборативной фильтрации можно отнести отсутствие необходимости знать что-либо о рекомендуемом предмете, так как рекомендации строятся на основе действий других пользователей.
-К минусам относится необходимость иметь данные об активности других пользователей, что называется ''холодный старт''.
+Пользовательский вектор имеет такую же размерность, что и вектор элемента.
 
-В свою очередь рекомендации на основе контента не требуют данных о действиях других пользователей, но необходимы признаки, которыми обладает рекомендуемый предмет, для построения вектора предпочтений пользователя.
+Поскольку теперь у нас есть векторы пользователей и элементов, мы можем найти между ними сходство:
 
-\subsection{Подходы}
-
-Для коллаборативной фильтрации используют метод kNN \cite{wiki:knn} и корреляцию Пирсона для нахождения похожих векторов.
-
-Для рекомендации на основе содержимого используются байесовский классификатор, решающие деревья, кластерный анализ.
-
-\section{Решение}
-
-В качестве решения предлагается комбинация подходов:
-
-\begin{figure}[h]
-	\includegraphics[width=0.5\textwidth]{./images/main.png}
-	\centering
-\end{figure}
-
-Подходы:
+\begin{equation}
+    r_{ui}=\cos (\theta)=\frac{\mathbf{u} \cdot \mathbf{v}}{\|\mathbf{u}\|\|\mathbf{v}\|}=\frac{\sum_{i=1}^{n} u_{i} v_{i}}{\sqrt{\sum_{i=1}^{n} u_{i}^{2}} \sqrt{\sum_{i=1}^{n} v_{i}^{2}}}
+\end{equation}
 
 \begin{itemize}
-	\item Content-based recommender
-	\item Collaborative recommender
-	\item Session recommender
-	      \begin{itemize}
-		      \item рекомендации на основе последних просмотренных элементов
-	      \end{itemize}
-	\item Boost recommender
-	      \begin{itemize}
-		      \item рекомендации на основе информации об элементе
-	      \end{itemize}
+    \item $r_{ui}$ -- рейтинг $user$ $item$
+    \item $u$ -- вектор $user$
+    \item $v$ -- вектор $item$
 \end{itemize}
 
-Каждый из сервисов уже был реализован, остаётся решить задачу оптимизации, то есть найти подходящие веса для каждого из подходов.
+Как было сказано ранее, фильтрация на основе содержимого имеет свой собственный вес воздействия, который вначале невелик (мы не хотим ограничивать доступ пользователя к контенту только потому, что он сделал несколько случайных щелчков), но он увеличивается по мере взаимодействия пользователя с системой и мы можем делать прогнозы относительно его категориальных предпочтений.
 
-\section{Дальнейшие планы}
+\subsection{Обзор компонентов ранжирования}
 
-В начале января планируется релиз приложение и масштабная пиар-компания, вследствие чего будет доступно очень много логов пользовательской активности.
+\subsubsection{Коллаборативная фильтрация}
 
-Собирая и используя эту информацию мы сможем оценить качество построенных моделей.
+Мы используем алгоритм SVD ++.
 
-Также планируется проводить A / B тестирование.
+Для рекомендаций по коллаборативной фильтрации у нас должна быть матрица пользовательских оценок, которая формируется из истории оценок.
+
+Мы используем модификацию Funk MF, которая факторизует матрицу рейтингов «пользователь-элемент» как произведение двух матриц меньшей размерности, первая имеет строку для каждого пользователя, а вторая имеет столбец для каждого элемента.
+Строка или столбец, связанные с конкретным пользователем или элементом, называются скрытыми признаками.
+
+\begin{equation}
+    r_{u i}=\sum_{f=0}^{n } H_{u, f} W_{f, i}
+\end{equation}
+
+Хотя Funk MF может обеспечить очень хорошее качество рекомендаций, но его способность использовать только явные числовые рейтинги в качестве взаимодействий между пользователями и элементами представляет собой ограничение.
+Современные рекомендательные системы должны использовать все доступные взаимодействия, как явные (например, числовые рейтинги), так и неявные (например, лайки, покупки, пропущенные, добавленные в закладки).
+С этой целью SVD ++ также был разработан с учетом неявных взаимодействий.
+По сравнению с Funk MF, SVD ++ также учитывает предвзятость пользователя и элемента.
+
+Прогнозируемый рейтинг, который пользователь $ u $ поставит элементу $ i $, вычисляется следующим образом:
+
+\begin{equation}
+    r_{u i}=\mu+b_{i}+b_{u}+\sum_{f=0}^{n} H_{u, f} W_{f, i}
+\end{equation}
+
+\subsubsection{Фильтрация по популярности}
+
+Для измерения популярности мы аггрегируем следующие данные: \hyperref[tab:show]{показы}, \hyperref[tab:view]{просмотры}, \hyperref[tab:emotion]{emoji}, \hyperref[tab:comment]{комментарии}.
+
+\begin{table}[h]
+    \centering
+    % \resizebox{\textwidth}{!}{
+    \begin{tabular}{ccccc}
+        \toprule
+
+        \emph{item\_id} & \emph{shows\_num} & \emph{views\_num} & \emph{emotions\_num} & \emph{comments\_num} \\\midrule
+
+        1 & 1043 & 231 & 52 & 7  \\
+        2 & 828  & 478 & 78 & 11 \\
+        3 & 163  & 25  & 5  & 0  \\
+        \vdots & \vdots & \vdots & \vdots & \vdots \\\bottomrule
+
+     \hline
+    \end{tabular}
+    % }
+
+    \caption{популярность элементов}
+    \label{tab:popularity}
+\end{table}
+
+Таким образом, когда мы применяем информацию о популярности новостных статей, мы можем выставлять им оценки по следующему алгоритму:
+
+\texttt{min-max} нормализуем \emph{shows\_num}, \emph{views\_num}, \emph{emotions\_num}, \emph{comments\_num}, затем делим на 4 (чтобы получить 1 как максимум после суммы) и суммируем все эти значения.
+
+\begin{equation}
+    r_i = \frac{\emph{shows\_num}}{4} + \frac{\emph{views\_num}}{4} + \frac{\emph{emotions\_num}}{4} + \frac{\emph{comments\_num}}{4} 
+\end{equation}
+
+\subsubsection{Фильтрация на основе содежимого (LDA)}
+
+Для фильтрации на основе содержимого мы должны каким-то образом векторизовать новости и представлять предпочтения пользователей через эти векторизованные новости.
+Мы будем использовать латентное распределение Дирихле (LDA), которое представляет собой генеративную статистическую модель, которая позволяет объяснять наборы наблюдений группами, которые объясняют, почему некоторые части данных похожи.
+Например, если наблюдения представляют собой слова, собранные в тексты, предполагается, что каждый документ представляет собой смесь небольшого количества тем и что присутствие каждого слова связано с одной из тем документа.
+
+Таким образом, мы можем векторизовать текст в зависимости от того, как каждый текст относится к каждой категории, от 0 до 1.
+Например, предположим, что у нас есть 3 темы, извлеченные с помощью модели LDA:
+
+\begin{table}[h]
+\centering
+% \resizebox{\textwidth}{!}{%
+\begin{tabular}{ccc}
+    \toprule
+\textit{topic} & \textit{word} & \textit{score} \\
+    \midrule
+1 & dollar & 0.3 \\
+1 & bank & 0.2 \\
+1 & money & 0.15 \\
+2 & sugar & 0.4 \\
+2 & cooking & 0.3 \\
+3 & IT & 0.25 \\
+3 & hacker & 0.2 \\
+\bottomrule
+\end{tabular}%
+% }
+\caption{темы слов}
+\label{tab:words_topics}
+\end{table}
+
+Допустим у нас текст: ``Hackers use mobile emulators to steal millions of dollars''. Его векторизованной формой будет $[0.3, 0, 0.2]$.
+
+Все вычисления проведены в соостветствии с \ref{content}.
+
+\subsubsection{Фильтрация на основе текущей сессии}
+
+Мы хотим мгновенно реагировать на действия пользователя, поэтому применяем фильтрацию на основе текущей сессии следующим образом:
+пытаемся найти новости, похожие на те, которые пользователь только что смотрел.
+
+\begin{equation}
+    r_{i}=\sum_{k=0}^{n} \textit{similarity}\{\textit{current\_item\_vector}, \textit{last\_viewed\_vector}_{k}\} \times \textit{weight}_k
+\end{equation}
+
+где $\textit{weight}_k$ это вес последней просмотренной новости. Вес тем больше, чем недавнее была просмотрена новость.
+
+\subsection{Гибридная система}
+
+Таким образом, наличие всей этой информации позволяет нам настраивать вклад каждой отдельной системы с помощью поиска по сетке.
+Таким образом, мы оптимизируем веса, которые влияют на каждую рекомендацию.
+Конкретные оптимизируемые метрики будут указаны ниже.
+
+% \section{Implementation}
+% \label{sec:implementation}
+
+\section{Оценка качества}
+\label{sec:evaluation}
+
+\subsection{Обзор}
+
+Для оценки у нас есть информация о том, какой список рекомендаций был дан каждому пользователю и каков источник рекомендации:
+
+    \begin{table}[h]
+        \centering
+        \resizebox{\textwidth}{!}{  
+        \begin{tabular}{cccc}
+            \toprule
+            \textit{user\_id} & \textit{recommendation\_list}       & \textit{content\_based\_filtering}  & \textit{collaborative\_filtering} \\
+            \midrule
+            2 & \{(2, 0.91), (1, 0.74), (3, 0.23)\} & \{(2, 0.45), (1, 0.54), (3, 0.08)\} & \{(2, 0.92), (1, 0.4), (3, 0.3)\}\\
+
+            1 & \{(3, 0.73), (1, 0.69), (2, 0.15)\} & \{(2, 0.6), (1, 0.44), (3, 0.04)\} & \{(2, 0.58), (1, 0.58), (3, 0.14)\}\\
+            \vdots & \vdots & \vdots & \vdots \\
+            \bottomrule
+        \end{tabular}%
+        }
+
+        \caption{логи выдачи рекомендаций}
+        \label{tab:recommendation_logs}
+    \end{table}
+
+    \begin{itemize}
+        \item \textit{recommendation\_list}: список рекомендаций, который состоит из пар (\textit{item\_id}, \textit{score})
+        \item \textit{content\_based\_filtering \& collaborative\_filtering}: источники рекомендаций
+    \end{itemize}
+
+  Поскольку мы используем взвешенную сумму наших рекомендаций, у нас есть уникальные значения весов для каждого пользователя:
+
+    \begin{table}[h]
+        \centering
+        \begin{tabular}{ccc}
+            \toprule
+            \textit{user\_id} & \textit{content\_based\_filtering} & \textit{collaborative\_filtering} \\
+            \midrule
+            1                 & 0.25                                  & 1                               \\
+            2                 & 1                               & 0.5                                \\
+            \vdots & \vdots & \vdots \\
+            \bottomrule
+            \end{tabular}%
+        \caption{значения весов}
+        \label{tab:boost_values}
+    \end{table}
+
+Чтобы проиллюстрировать подсчет оценок рекомендаций, взгляните на 1-ю строку таблицы \hyperref [tab:advice_logs] {журналы рекомендаций}.
+Мы видим рекомендацию $ r = \{(2, 0.91), (1, 0.74), (3, 0.23) \} $ для пользователя $ u = 2 $, поэтому мы должны найти значения повышения для этого конкретного пользователя в таблице \hyperref [tab:boost_values] {boost values}: усиление контентной фильтрации $ b_{cbf} $ для $ u $ составляет $ 1 $, усиление совместной фильтрации $ b_{cf} $ составляет $ 0,5 $.
+Итоговая оценка рассчитывается следующим образом: $ 0,91 = 0,45 * b_{cbf} + 0,92 * b_{cf} = 0,45 * 1 + 0,92 * 0,5 = 0,45 + 0,46 $.
+
+    Также у нас есть информация о \hyperref [tab:show] {shows} и \hyperref [tab:view] {views}, поэтому мы можем отслеживать, какой элемент в списке рекомендаций был нажат, а какой элемент был пропущен, поэтому в зависимости от этого информация, которую мы можем отслеживать, понравился ли элемент пользователю или нет.
+Например, рассмотрим ситуацию, когда пользователь щелкнул 4-ю рекомендацию из списка рекомендаций.
+Это означает, что первые 3 рекомендации плохие, и мы должны штрафовать алгоритмы, которые отдавали приоритет этим элементам, учитывая тот факт, что 1-й элемент является худшим:
+
+\begin{equation}
+    \textit{penalty} = \frac{1}{\textit{num\_of\_item\_in\_rank}}
+\end{equation}
+
+% \subsection{Online}
+
+% \subsection{Offline}
+
+% \section{Further research}
+% \label{sec:further}
+
+\section{Итоги}
+\label{sec:summary}
+
+Система рекомендаций получила широкое распространение в разных сферах.
+Коллаборативная фильтрация фокусируется на рейтинге, игнорируя особенности самих элементов.
+Чтобы лучше оценивать предпочтения клиентов в отношении новостей, мы используем модель LDA для рассчёта предпочтения клиентов по новостным темам.
+
+Чтобы прогнозировать рейтинг по новостям, мы принимаем во внимание схожесть клиентов и корреляцию между покупателями и новостями.
+Эксперимент показывает, что наш гибридный метод рекомендаций, основанный на характеристиках, лучше работает в нашем приложении для социальных сетей.
+
+Мы предлагаем новый гибридный метод рекомендаций, основанный на функциях повышения качества выдачи.
+
+Результаты показывают, что сочетание разных подходов приводит к увеличению вовлеченности пользователей.
+До внедрения рекомендательной системы пользователь тратил на вкладку новостей около 2 минут.
+Теперь, когда мы применили гибридную рекомендательную систему, пользователи проводят на вкладке новостей в среднем 5 минут.
 
